\section{Early consistency models}
\label{sec:early_models}

 \textbf{1:1-model}: The very early models contain nothing but a base 
table on one node (i.e. a standalone database) and a view table on 
another node (i.e. the data warehouse) \cite{zhuge:view}. The base table 
is denoted as $B$ and the view table as $V$. Once operations $t_1..t_n 
\in T$ are applied to the base table $B$, the base table changes its 
state. This leads to different \textit{base states} that are denoted as 
$B_0..B_i..B_f$. The initial state of the base table is $B_0$. Every 
operation on the base table produces a new base state until the final 
state $B_f$ is reached. The operator $\leq$ is used to express the 
relation between states, e.g. $B_i \leq B_j$. A state $B_j$ is said to 
be greater than a state $B_i$ if $B_j$ contains all operations of $B_i$ 
plus a number of subsequent operations. Formally, we demand $(\forall t 
\in B_i)\;t \in B_j$ and $(\exists t_k \in B_j)(\forall t_l \in B_i)\;k 
> l $. The view table in the data warehouse also consists of a series of 
\textit{view states} $V_0..V_j..V_f$. Consistency is analysed by 
comparing the system's view states with the theoretical correct view 
states, derived from the appropriate base state. We derive the correct 
view state by applying function $View(B)$. Thus, we say a view state 
corresponds to a base state when $V_j=View(B_i)$. 


\textbf{1:N model}: In a next step, the consistency model was extended 
to describe view maintenance in a multi-source data 
warehouse\cite{zhuge:strobe}. Here, the data warehouse is still located 
on one node, but the base table information is consumed from $n$ 
different source nodes. Operations on a single source are serializable 
and executed in sequence. However, operations from different sources 
doesn't need to preserve ordering. The base table states are defined as 
$B_0..B_i..B_f$. The definition $\leq$ can be applied between the 
produced states, again. But, in order to match the relaxed ordering 
requirements, we have to extend the declaration from before to $(\exists 
t_k \in B_j)(\forall t_l \in B_i)\;k > l\;\land\;s(t_k)=s(t_l)$. 
Function $s()$ delivers, when applied to an operation, the source from 
which the operation originated. However, two operations on different 
sources are processed in parallel, allowing two possible intermediary 
base states $B_i$ and $B_j$. $B_i$ does not contain all operations of 
$B_j$ and vice versa. Thus, $B_i$ and $B_j$ are neither $\leq$ nor 
$\geq$. We denote this relation as $B_i \parallel B_j$. As already 
stated, this extended model represents a relaxation to the original 
model. We do not claim global serializability of operations any more. 

